---
title: "gt"
output: html_document
---


```{r, include=F}
options(knitr.duplicate.label = "allow")
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = F, echo = F)
library(flipbookr)
library(tidyverse)
```


---

## a taste of gt {.tabset .tabset-fade .tabset-pills}


### example prep

Using the exibble data that comes with gt, but I like a more more concrete variable names so doing some prep here:

```{r gt_prep, include = F}
library(gt)
gt::exibble %>% 
  rename(fruit = char,
         team = fctr,
         weight = num,
         holiday = date,
         schedule = time,
         appointment = datetime, 
         price = currency,
         id = row,
         club = group) ->
example; example
```


```{r}
embed_flipbook("gt_prep", 
               font_size = 75,
               use_share_again = T,
               use_embed_xaringan = T)
```

---

### gt overview


```{r gt2}
library(gt)
example %>%
  .[1:5, ] %>% 
  gt() %>%
  fmt_number(columns = vars(weight), 
             decimals = 2) %>%
  fmt_date(columns = vars(holiday), 
           date_style = 1) %>% 
  fmt_time(columns = vars(schedule), 
           time_style = 1) %>%
  fmt_datetime(columns = vars(appointment), 
               date_style = 6, 
               time_style = 4) %>%
  fmt_currency(columns = vars(price), 
               currency = "EUR") %>%
  tab_header(
    title = md("This is the `exibble` dataset in **gt**"), 
    subtitle = "It is one of six datasets in the package"
  ) %>%
  tab_source_note(source_note = 
                    md("More information is available at `?exibble`.")) %>% 
  tab_spanner(label = "Dates and Times", 
              columns = matches("schedule|appointment"))
```


```{r}
embed_flipbook("gt2", 
               font_size = 75,
               use_share_again = T,
               use_embed_xaringan = T)
```

## formating {.tabset .tabset-fade .tabset-pills}



### Formatted values: time




```{r gt_time, include=F}
library(gt)  
example %>%  
  .[1:5, 5] %>%  
  gt() %>%  
  fmt_time(columns = vars(schedule), time_style = "hms") %>% #ROTATE
  fmt_time(columns = vars(schedule), time_style = "hm") %>% #ROTATE
  fmt_time(columns = vars(schedule), time_style = "hms_p") %>% #ROTATE
  fmt_time(columns = vars(schedule), time_style = "hm_p") %>% #ROTATE
  fmt_time(columns = vars(schedule), time_style = "h_p") #ROTATE
```

```{r}
embed_flipbook("gt_time", 
               break_type = "rotate", 
               font_size = 90, 
               widths = c(2,1),
               use_share_again = T,
               use_embed_xaringan = T)
```

### Formatted values: date

---


```{r gt_date, include=F}
library(gt)  
example %>%  
  .[1:5, 4] %>%  
  gt() %>%  
  fmt_date(columns = vars(holiday), 
           date_style = "iso")
```

```{r}
embed_flipbook(chunk_name = "gt_date", 
               break_type = "replace", 
               replace = "iso",
               font_size = 90, widths = c(2,1),
               replacements = c("iso","wday_month_day_year", "wd_m_day_year", 
                                "wday_day_month_year", "month_day_year", 
                                "m_day_year", "day_m_year", "day_month_year", 
                                "day_month", "year", "month", "day", 
                                "year.mn.day", "y.mn.day"),
               use_share_again = T,
               use_embed_xaringan = T)
```



### Formatted values: currencies


---


```{r gt_currency, include=F}
library(gt)  
example %>%  
  .[1:8, 7] %>%  
  gt() %>%  
  fmt_currency(columns = vars(price), 
               currency = "USD")
```



```{r}
embed_flipbook("gt_currency",
               break_type = "replace",
               replace = "USD",
               replacements = c(
                 "AED", "AFN", "BAM", "BBD", "BDT", "BGN", "BHD", "TMT", "TND", 
                 "TOP", "USD", "USN", "USS", "UYI", "UYU", "UZS", "VEF", "VND", 
                 "VUV", "XPD", "XPF", "XPT", "XSU", "XUA", "YER", "ZAR", "ZMW"
               ),
               font_size = 90, 
               widths = c(2,1),
               use_share_again = T,
               use_embed_xaringan = T)
```


## Prepping a pizza menu in four parts {.tabset .tabset-fade .tabset-pills}

### part 1

```{r gt_pizza}
library(tidyverse)
library(paletteer)
library(gt)
pizzaplace %>%
  mutate(type = case_when(
    type == "chicken" ~ "chicken (pizzas with chicken as a major ingredient)",
    type == "classic" ~ "classic (classical pizzas)",
    type == "supreme" ~ "supreme (pizzas that try a little harder)",
    type == "veggie" ~ "chicken (pizzas without any meats whatsoever)",
  )) %>%
  mutate(size = factor(size, 
                       levels = c("S", "M", "L", 
                                  "XL", "XXL"))) %>%
  dplyr::group_by(type, size) %>%
  dplyr::summarize(
    sold = n(),
    income = sum(price)
  ) %>% 
  # table is kind of large, hard to make sense of
  dplyr::filter(size != "S" & 
                  size != "M" & 
                  size != "XL") ->
pizza_prep; pizza_prep
```


```{r}
embed_flipbook("gt_pizza",
               font_size = 90, widths = c(1,1),
               use_share_again = T,
               use_embed_xaringan = T)
```




---


### part 2

```{r gt_pizza_prep}
pizza_prep %>% 
  gt(rowname_col = "size") %>%
  tab_header(title = 
               md("&#127829; Pizzas Sold in 2015 &#127829;")) %>%
  fmt_number(
    columns = vars(sold),
    decimals = 0,
    use_seps = TRUE
  ) %>%
  fmt_currency(
    columns = vars(income),
    currency = "USD"
  ) %>%
  cols_align(align = "right", columns = TRUE) ->
pizza2; pizza2
```


```{r}
embed_flipbook("gt_pizza_prep",
               font_size = 90,
               use_share_again = T,
               use_embed_xaringan = T)
```



---

### part 3

```{r gt_pizza1}
pizza2 %>% 
  data_color(
    columns = vars(sold, income),
    colors = scales::col_numeric(
      palette = paletteer::paletteer_d(
        palette = "ggsci::red_material"
      ) %>% as.character(),
      domain = NULL
    ),
    alpha = 0.8
  ) %>%
  summary_rows(
    groups = TRUE,
    columns = vars(sold),
    fns = list(TOTAL = "sum"),
    formatter = fmt_number,
    decimals = 0,
    use_seps = TRUE
  ) %>%
  summary_rows(
    groups = TRUE,
    columns = vars(income),
    fns = list(TOTAL = "sum"),
    formatter = fmt_currency,
    currency = "USD"
  ) %>%
  grand_summary_rows(
    columns = vars(sold),
    fns = list(`GRAND TOTAL` = "sum"),
    formatter = fmt_number,
    decimals = 0,
    use_seps = TRUE
  )  ->
pizza3; pizza3
```

```{r}
embed_flipbook("gt_pizza1", 
               font_size = 75,
               use_share_again = T,
               use_embed_xaringan = T)
```


### part 4

---

```{r gt_pizza2}
pizza3 %>%
  grand_summary_rows(
    columns = vars(income),
    fns = list(`GRAND TOTAL` = "sum"),
    formatter = fmt_currency,
    currency = "USD"
  ) %>%
  tab_options(table.font.size = "tiny") %>% 
  tab_footnote(
    footnote = "The pizza category with the highest total sales.",
    locations = cells_row_groups("classic (classical pizzas)")
  ) %>%
  tab_footnote(
    footnote = md("Custom sizes for **The Greek** pizza."),
    locations = cells_stub("XXL")
  ) %>%
  tab_footnote(
    footnote = md("This is a new record. Truly, 2015 was a **great** year for the `pizzaplace`."),
    locations = cells_grand_summary(columns = vars(sold))
  ) ->
pizza_tab; pizza_tab
```  
  
  

```{r}
embed_flipbook("gt_pizza2", 
               font_size = 75,
               use_share_again = T,
               use_embed_xaringan = T)
```




```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 200%}

@media print {
  .has-continuation {
    display: block;
  }
}
```

