---
title: "Named colors in R"
author: "Evangeline Reynolds"
date: "4/25/2020"
output: html_document
---



```{r, include=F}
options(knitr.duplicate.label = "allow")
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = F, echo = F)
library(flipbookr)
library(tidyverse)
```


*Base R contains 657 different color names that can be used to get colors; those names can be retrieved with the `colors()` function.  There are some names that refer to the same color like "gray" and "grey", and "bisque" and "bisque1".  Most names and corresponding colors are listed here alphabetically, except for greys/grays which are show at the bottom.  A few named colors are not reproduced here.*  


```{r named_colors_colorful}
library(tidyverse)

my_colors <- colors()[!str_detect(colors(), "^grey|^gray|indian|peru|navajo")]
grays <- colors()[str_detect(colors(), "^grey|^gray")]

num_columns <- 7
num_colors <- length(my_colors)
start_color <- 1
end_color <- start_color + 
  num_colors - 1
the_colors <- my_colors[start_color:end_color]


the_colors %>% 
  matrix(ncol = num_columns, byrow = F, ) %>% 
  as_tibble() %>% 
  mutate_at(.vars = 1:num_columns, 
            funs(factor(., 
                        levels = 
                          rev(the_colors), 
                        ordered = T))) %>% 
  gt::gt() %>% 
  gt::data_color(columns = 1:num_columns,
                 colors = scales::col_factor(
                   palette = the_colors,
                   domain = NULL,
                   reverse = TRUE)
                 ) %>%
  gt::tab_options(
    column_labels.hidden = TRUE
  )
```

```{r}
embed_flipbook("named_colors_colorful",
               use_share_again = T,
               use_embed_xaringan = T,
               font_size = 80)
```



```{r named_colors_greys}
num_columns <- 12
num_colors <- length(grays)
start_color <- 1
end_color <- start_color + num_colors - 1
the_colors <- grays[start_color:end_color]


the_colors %>% 
  matrix(ncol = num_columns, byrow = F, ) %>% 
  as_tibble() %>% 
  mutate_at(.vars = 1:num_columns, 
            funs(factor(., 
                        levels = rev(the_colors), 
                        ordered = T))) %>% 
  gt::gt() %>% 
  gt::data_color(columns = 1:num_columns,
                 colors = scales::col_factor(
                   palette = the_colors,
                   domain = NULL,
                   reverse = TRUE)
                 ) %>%
  gt::tab_options(
    column_labels.hidden = TRUE
  )
```

```{r}
embed_flipbook("named_colors_greys",
               use_share_again = T,
               use_embed_xaringan = T,
               font_size = 80)
```

