---
title: "Table formatting!"
subtitle: "With flipbookr and xaringan"
author: "Gina Reynolds, December 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = F, dev = "svg", fig.ext = "svg")
library(flipbookr)
library(tidyverse)
```








---



`r chunk_reveal("tables", break_type = "rotate", title = "### Some of our formatting options")`

```{r tables, eval = F, echo = F}
gapminder::gapminder %>% 
  filter(year == 2002) %>% 
  tibble() #ROTATE
  data.frame() #ROTATE
  data.table::data.table() #ROTATE
  flextable::flextable() #ROTATE
  knitr::kable() #ROTATE
  gt::gt() #ROTATE
  DT::datatable() #ROTATE
```



---

# list column

---

`r chunk_reveal("listcolumn", break_type = "rotate", title = "### What they do with the list column..." )`

```{r listcolumn, eval = F, echo = F}
tibble(x = c(1,2,3,4,6),
       y = 1:5,
       z = list(1:2, 4:5, 8, NA, c(1,7,9,0))) %>% 
  tibble() #ROTATE
  data.frame() #ROTATE
  data.table::data.table() #ROTATE
  knitr::kable() #ROTATE
  gt::gt() #ROTATE
  DT::datatable() #ROTATE
```

---

# group

---

`r chunk_reveal("group", break_type = 1, title = "### What they do with grouped data..." )`

```{r group, eval = F, echo = F}
gapminder::gapminder %>%
  group_by(country) %>% 
  filter(year > 1990) %>% 
  filter(continent == "Americas") %>% 
  select(-continent) %>% 
  gt::gt()
```



---

`r chunk_reveal("margins_table_tidy", title = "### using tidyverse 'verbs' to create pop count by continent in 2002 with percentages and totals")`



```{r margins_table_tidy, include=FALSE}
gapminder::gapminder %>% 
  filter(year == 2002) %>% 
  group_by(continent) %>% 
  summarise(population = sum(pop)) %>% 
  mutate(percent = 100 * round(prop.table(population), 3) ) ->
base
  
base %>% 
  summarise(percent = sum(percent), 
            population = sum(population),
            continent = "Totals") ->
totals

bind_rows(base, totals) %>% 
  mutate(`Population (millions)` = population/1000000) %>% 
  mutate(percent = round(percent, 2)) %>% 
  rename(` ` = continent) %>% 
  select(` `, `Population (millions)`, percent)
```


---

# That was a lot of code... What other options do we have?


---

`r chunk_reveal("prop.table", title = "### What about the very classic table and prop.table?")`

```{r prop.table, include = F}
gapminder::gapminder %>% 
  filter(year == 2002) %>% 
  pull(continent) %>% 
  table() %>% 
  prop.table() %>% 
  `*`(100) %>% 
  round(1) %>% 
  paste0("%")
```

---


```{r}
library(gtsummary)



```



---


We are working with maintainers of many of the R community’s most popular R packages for building tables, including Yihui Xie of DT, Rich Iannone of gt, Greg Lin of reactable, David Gohel of flextable, David Hugh-Jones of huxtable , and Hao Zhu of kableExtra. Many of these maintainers will help review submissions built with their packages.


---



```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 80%}

@media print {
  .has-continuation {
    display: block;
  }
}
```
